---
title: "Assignment 4"
author: "Saket Choudhary"
date: "March 20, 2016"
output: html_document
---

## Problem 1

```{r}
x1 <- rexp(n=100)
m.x1 <- mean(x1)
m.x1
```

```{r}
sd.x1 <- sd(x1)
sd.x1
```

Mean: `r m.x1`

Standard Deviation: `r sd.x1`

## Problem 2
```{r}
x0.1 <- rexp(n=100, rate=0.1)
x0.5 <- rexp(n=100, rate=0.5)
x5 <- rexp(n=100, rate=5)
x10 <- rexp(n=100, rate=10)
```
`x0.1`: Mean = `r mean(x0.1)` SD  = `r sd(x0.1)`

`x0.5`: Mean = `r mean(x0.5)` SD  = `r sd(x0.5)`

`x5`: Mean = `r mean(x5)` SD  = `r sd(x5)`

`x10`: Mean = `r mean(x10)` SD  = `r sd(x10)`


## Problem 3

### Problem 3.a

```{r}
plot(c(0.1,0.5,1,5,10), c(mean(x0.1), mean(x0.5), mean(x1), mean(x5), mean(x10)), pch=2, xlab = 'Rate', ylab='Mean', main='Mean vs Rate')
```


### Problem 3.b

```{r}
plot(c(0.1,0.5,1,5,10), c(sd(x0.1), sd(x0.5), sd(x1), sd(x5), sd(x10)), pch=2, xlab = 'Rate', ylab='SD', main='SD vs Rate')
```

### Problem 3.c

```{r}
plot(c(mean(x0.1), mean(x0.5), mean(x1), mean(x5), mean(x10)), c(sd(x0.1), sd(x0.5), sd(x1), sd(x5), sd(x10)), pch=2, xlab = 'SD', ylab='Mean', main='Mean vs SD')
```

$E[X] = \frac{1}{\lambda}$ and $Var[X] = \frac{1}{\lambda^2}$ hence $E[X] vs \sqrt{Var[X]}$ follows a linear trend

## Problem 4

```{r}
y <- rexp(n=1100000)
```

Mean of $y$: `r mean(y)`
SD of $y$: `r sd(y)`

## Problem 5

```{r}
hist(y)
```

$y$ is the PDF of exponential distribution and hence matches the following distribution $e^{-x}$

```{r}
h <- hist(y, plot = FALSE)
ylim <- range(0, h$density, dexp(0))
hist(y, freq = FALSE, ylim = ylim)
curve(dexp, y, add=TRUE)
```

```{r}
plot(seq(0,15,0.1), exp(-seq(0,15,0.1)), main='$e^{-x}$ vs $x$', xlab = '$x$', ylab='$y$')
```


## Problem 6
```{r}
y.mat  <- matrix(y, nrow=1100,ncol=1000)
```

## Problem 7

```{r}
mean(y.mat[,371])
```

## Problem 8

```{r}
col.m <- colMeans(y.mat)
h <- hist(col.m, plot = FALSE)
ylim <- range(0, h$density, dnorm(0))
hist(col.m, freq = FALSE, ylim = ylim)
curve(dnorm, col.m, add=TRUE)
```

The shape of the column means soes not match that of Problem 2 because of central limit theorem, which states 
that the distribution of mean of large number of iterants(columns in this case) is approximately normal

### Problem 9

```{r}
mean(y.mat[y.mat>1])
```

#Part 2

```{r}
temp.data <- read.csv("Temperature.csv", header=TRUE)
temp.data$DateNr <- as.Date(temp.data$DateNr, format='%m/%d/%Y')
month <- format(temp.data$DateNr, '%m')
temp.data$month <- month
uniq.months <- sort(unique(month))
splitted.data <- split(temp.data, temp.data$month)

```
